---
title: "Data Processing and Analysis Project"
output:
  html_document:
    df_print: paged
  html_notebook: default
---

This is a final project. Predictable, I've decided to analyze [data](https://www.kaggle.com/sudalairajkumar/novel-corona-virus-2019-dataset) connected with new *coronavirus*.

Importing all needed packages:
```{r}
library(ggplot2)
library(psych)
library(gtable)
library(grid)
library(gridExtra)
library(stringr)
```


Loading the data to dataframe and checking first five rows:
```{r}
df <- read.csv('~/Documents/se/R/novel-corona-virus-2019-dataset/covid_19_data.csv')
head(df)
```

As we can see, there are 8 columns with basic info about COVID-19 spreading. Interesting columns for this project: *Country.Region*, *ObservationDate*, *Confirmed*, *Deaths*, *Recovered*.

Changing the type of date column:
```{r}
df$ODate <- as.Date(df$ObservationDate, format = "%m/%d/%y")
```

The first distribution which was interested to investigate is *Confirmation of coronavirus*. Plots below illustrates that the form of confirmation distribution is sammilar to exponent function curve:
```{r}
confirmed_by_days <- aggregate(Confirmed ~ ODate, df, sum)
exp_x <- seq(0, 20, length.out = 1000)
exp_data <- data.frame(x=exp_x, y = exp(exp_x))

p1 <- ggplot(confirmed_by_days, aes(x=ODate, y=Confirmed)) +
  geom_line(size=1.5) + 
  xlab("Date") + 
  ylab("Number of confirmed cases") +
  ggtitle("Distribution of \n coronavirus confirmation") +
  theme(plot.title = element_text(face='bold', 
                                  colour='black', size=18, hjust = 0.5)) +
  scale_x_date(date_labels = "%b %d", date_breaks = "30 days")
p2 <- ggplot(data = exp_data, aes(x=x, y=y)) +
  geom_line(size=1.5) +
  xlab("x") + 
  ylab("exp(x)") +
  ggtitle("Distribution of \n exponential function") +
  theme(plot.title = element_text(face='bold', colour='black', size=18, hjust = 0.5)) 

grid.arrange(p1, p2, nrow = 1)
```

But, I think, it's not a reason for a panic, because we (humanity) can change the constant inside the exponent function and this distribution will turn into *Logistic curve*. 


The sad plot is a distribution deaths by days. Spoiler: it depends on a number of confirmed issues, so it's obvious that the curve will be simillar...
```{r}
deaths_by_days <- aggregate(Deaths ~ ODate, df, sum)

ggplot(deaths_by_days, aes(x=ODate, y=Deaths)) +
  geom_line(size=1.5) + 
  xlab("Date") + 
  ylab("Number of deaths") +
  ggtitle("Distribution of \n death numbers") +
  theme(plot.title = element_text(face='bold', 
                                  colour='black', size=18, hjust = 0.5)) +
  scale_x_date(date_labels = "%b %d", date_breaks = "30 days")
```
```{r}
ggplot(aggregate(russia$Deaths ~ russia$ODate, russia, FUN = sum), 
       aes(x=russia$ODate, y=russia$Deaths)) +
  geom_line()
```

Time for good news: there is distribution of recovering (guess what, it depends on number of confirmed!)
```{r}
recovered_by_days <- aggregate(Recovered ~ ODate, df, sum)

ggplot(recovered_by_days, aes(x=ODate, y=Recovered)) +
  geom_line(size=1.5) + 
  xlab("Date") + 
  ylab("Number of recovered") +
  ggtitle("Distribution of \n healed numbers") +
  theme(plot.title = element_text(face='bold', 
                                  colour='black', size=18, hjust = 0.5)) +
  scale_x_date(date_labels = "%b %d", date_breaks = "30 days")
```

To estimate the real mortality situation we should rate the $\frac{number of deaths}{number of confirmed}$:
```{r}
mortality <- data.frame(x=df$ODate, y = df$Deaths/df$Confirmed)
mortality <- mortality[mortality$x>"2020-03-01", ]
mortality <- mortality[complete.cases(mortality), ]

russia = subset(df, df$Country.Region == 'Russia')
r_mortality <- data.frame(x=russia$ODate, y = russia$Deaths/russia$Confirmed)
r_mortality <- r_mortality[r_mortality$x>"2020-03-01", ]
r_mortality <- r_mortality[complete.cases(r_mortality), ]

colors <- c("Russia" = "red", "ROW" = "blue")
ggplot(aggregate(y ~ x, r_mortality, mean), 
       aes(x=x, y=y, color='Russia')) +
  geom_line(size=1.5) +
  geom_line(data = aggregate(y ~ x, mortality, mean),
            aes(x = x, y = y, color = 'ROW'), size=1.5) +
  labs(x = 'Date',
    y = 'Mortality Rate, %',
    title = "Distribution of coronavirus mortality rate \n over Russia and the rest of the world",
    color = 'Legend') +
  theme(plot.title = element_text(face='bold', 
                                  colour='black', size=18, hjust = 0.5)) +
  scale_x_date(date_labels = "%b %d", date_breaks = "10 days") + 
  scale_y_continuous(labels = scales::percent) +
  scale_color_manual(values = colors)

```

Yes, there is a small growth, but the share of deaths is under 1% in Russia (or may be we don't know something...)

Speaking about Russia, let's check a summary:
```{r}
describe(russia[, c(6, 7, 8)], skew=F)
```
The maximum number of deaths is 615 over all regions.


Loading different data about COVID. Breaking down to confirmed issues:
```{r}
df = read.csv('~/Documents/se/R/novel-corona-virus-2019-dataset/COVID19_open_line_list.csv', 
              na.strings=c("","N/A", "NA"))
head(df)
```
There are interesting columns for this project: *age*, *sex*

Counting some statistic about genders:
```{r}
sex <- str_replace(df$sex, 'Female', 'female')
sex <- str_replace(sex, 'Male', 'male')
sex <- na.omit(sex)
sex <- sex [! sex == 4000]
table(sex)
```

Visualizing the summary:
```{r}
ggplot(data.frame(sex), aes(x=factor(1), fill=sex)) +
  geom_bar(width = 1) +
  coord_polar("y") +
  labs(title = "Gender ration over confirmed issue") +
  theme(plot.title = element_text(face='bold', 
                                  colour='black', size=18, hjust = 0.5))
```

Counting some stat about ages distribution:
```{r warning=FALSE}
ages <- as.numeric(df$age)
ages[is.na(ages)] <- mean(ages, na.rm=TRUE)
ages <- round(ages)
lbl <- c("0-10", "10-20", "20-30", "30-40", "40-50", "50-60", "60-70", 
"70-80", "80-90", "90-100")
ages_cat <- as.factor(cut(ages, breaks = seq(0,100,10), labels = lbl))

ggplot(data.frame(x=ages_cat), aes(x=x))+
  geom_bar(stat="count", width=0.7, fill="#56B4E9", na.rm = T) +
  theme_minimal() +
  labs(x = 'Age',
    y = 'Number of people in group',
    title = "Ages groups over confirmed issue") +
  theme(plot.title = element_text(face='bold', 
                                  colour='black', size=18, hjust = 0.5))
```

~~There is some bag in R, because as we can see there is no NA issues((~~

```{r}
table(ages_cat)
```

There are a lot of researches about COVID-19 in the Internet. The purpose of this Notebook is *getting experience* with data processing and visualizing via R language.

_Ideas_: deep into geographical data from the second dataset, look through sympthoms from the second dataset, aggregate data by location from the first dataset.



